---
title: "wine-quality"
author: "Inna Galat"
date: "9 July 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
ptStart <- proc.time()

library(data.table)
library(ggplot2)
library(corrplot)

projpath <- file.path("~","Github", "wine-quality-r")
datapath <- file.path(projpath,"data")
```

## Preface

GOAL: 
- to understand which wine properties affect the wine quality according to the expert valuation.
The response variable shows discreet categories between 0 (the worst) to 10 (the best).
This problem can be approached using regression or classification models. 

### 1. Reading Datasets

```{r datasets-reading}

# initialise wine as empty list
wine <- list()
for (filename in c("red", "white")) {
  file <- file.path(datapath,paste0("winequality-", filename,".csv"))
  wine[[filename]] <- read.table(file, sep=";", as.is=TRUE, header=TRUE)
}

summary(wine)
head(wine[[1]],3)
head(wine[[2]],3)
summary(wine$red)
summary(wine$white)


```

```{r corr}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor=1, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    
    r <- cor(x,y,method="spearman")
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    text(0.5, 0.5, txt, cex = cex.cor * (1+abs(r)))
}

pairs(wine[[1]], upper.panel=panel.cor)
pairs(wine[[2]], upper.panel=panel.cor)

```



## Including Plots

You can also embed plots, for example:

```{r plot, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
