---
title: "wine-quality"
author: "Inna Galat"
date: "9 July 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
ptStart <- proc.time()

library(data.table)
library(ggplot2)
library(corrplot)

projpath <- file.path("~","Github", "wine-quality-r")
datapath <- file.path(projpath,"data")
```

## Preface

GOAL: 
- to understand which wine properties affect the wine quality according to the expert valuation.
The response variable shows discreet categories between 0 (the worst) to 10 (the best).
This problem can be approached using regression or classification models. 

### 1. Reading Datasets

```{r datasets-reading}

# initialise wine as empty list
wine <- list()
for (filename in c("red", "white")) {
  file <- file.path(datapath,paste0("winequality-", filename,".csv"))
  wine[[filename]] <- read.table(file, sep=";", as.is=TRUE, header=TRUE)
}

summary(wine)
names(wine)
head(wine[[1]],3)
head(wine[[2]],3)
summary(wine$red)
summary(wine$white)


```

```{r corr}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor=1, ...)
{
    usr <- par("usr"); 
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    
    r <- cor(x,y,method="spearman")
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    text(0.5, 0.5, txt, cex = cex.cor * (1+abs(r)))

}

pairs(wine[[1]], upper.panel=panel.cor)
pairs(wine[[2]], upper.panel=panel.cor)
par(mfrow=c(1,1))

```


### Outliers detection and treatment
```{r outliers}

plot(wine[[1]]$total.sulfur.dioxide)
plot(wine[[1]]$chlorides)

plot(wine[[2]]$free.sulfur.dioxide)
plot(wine[[2]]$residual.sugar)
plot(wine[[2]]$density)
```
Outliers: 
Red Wine:
- total.sulfur.dioxide > 200
- chlorides > 0.5

White Wine:
free.sulfur.dioxide > 200
residual.sugar > 40
density > 1.05

Ways to treat outliers:
- to subset data by removing outliers
- to apply log-transformation, so the effect of large values on the distribution is subsided 
- to apply scale()


###Plotting distribution 

```{r distribution_plot}

density_plot <- function(j) { 
  old.par <- par(mfrow=c(3,4),ps=16)
  wine_obj <- wine[[j]]
  for (i in names(wine_obj)){
    # print(i)
    plot(density(scale(wine_obj[,i])), main=paste(j,":",i))
    wine_obj[,i][wine_obj[,i]==0] <- min(wine_obj[,i][wine_obj[,i]>0])/2
    logdensity <- density(scale(log(wine_obj[,i])))
    points(logdensity$x, logdensity$y, type="l", col=2)
    legend("topright",c("ori","log"),col=1:2,text.col=1:2,lty=2, cex=0.70)
  }
}
  

density_plot("red")
density_plot("white")
```

### Outliers Removal
```{r outliers_removal}

wine[[1]] <- wine[[1]][wine[[1]]$total.sulfur.dioxide<200 & wine[[1]]$chlorides<0.5, ]

wine[[2]] <- wine[[2]][wine[[2]]$free.sulfur.dioxide<200 & wine[[2]]$residual.sugar<40 & wine[[2]]$density<1.05, ]

density_plot("red")
density_plot("white")

```
